\textbf{ZKLoRA: Efficient Zero-Knowledge Proofs for LoRA Verification}

### 2025-01-17 paper/main.tex
updated tikz figure; 2 figures -> 1 table

We benchmarked \texttt{ZKLoRA} across several LLMs and smaller models with different numbers of LoRA modules. 
Our central question is how \emph{total} verification, proof generation, and settings times grow with \textbf{the number of LoRA modules} (\#LoRA), while also considering each LoRA's average parameter size. 
Figures~\ref{fig:verify_vs_numlora}, \ref{fig:proof_vs_numlora}, and \ref{fig:settings_vs_numlora} visualize this trade-off:
\begin{itemize}
    \item The x-axis is \textbf{the number of LoRA modules} for each model setup,
    \item The y-axis is one of: \textit{total verification time}, \textit{total proof generation time}, or \textit{total settings time},
    \item \textbf{Marker size} is proportional to the average LoRA size (in parameters).
    \item The legend indicates each model’s average LoRA size with a label \textit{“Avg LoRA Size=xxK”}.
\end{itemize}

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.65\linewidth]{figs/fig_verify_vs_numlora.pdf}
    \caption{Total verification time (seconds) vs. number of LoRA modules, with dot size reflecting average LoRA size.}
    \label{fig:verify_vs_numlora}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.65\linewidth]{figs/fig_proof_vs_numlora.pdf}
    \caption{Total proof generation time (seconds) vs. number of LoRA modules, with dot size reflecting average LoRA size.}
    \label{fig:proof_vs_numlora}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.65\linewidth]{figs/fig_settings_vs_numlora.pdf}
    \caption{Total settings time (seconds) vs. number of LoRA modules, with dot size reflecting average LoRA size.}
    \label{fig:settings_vs_numlora}
\end{figure}

@@

Figures~\ref{fig:proof_vs_numlora} and \ref{fig:settings_vs_numlora} similarly show that \emph{proof generation} and \emph{settings time} scale with the number of LoRA modules, but also depend heavily on each module’s average size (indicated by larger or smaller dot size). These two steps (proof generation on the LoRA contributor’s side and cryptographic circuit setup) can become more expensive, yet remain feasible in decentralized settings or paid contract relationships. The \emph{Base Model User}, meanwhile, benefits from the relatively short verification overhead.

@@

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
    font=\small,
    >=latex,
    thick,
    block/.style={
      draw,
      rounded corners,
      align=center,
      minimum width=1.65cm,  % Decreased from 1.8cm to 1.65cm
      minimum height=0.9cm,
      fill=gray!10  % Light gray background for regular blocks
    },
    dashedblock/.style={
      draw,
      dashed,
      rounded corners,
      align=center,
      font=\footnotesize,
      fill=orange!10,  % Light orange background for LoRA block
      minimum width=1.65cm,  % Decreased to match other blocks
      minimum height=0.9cm
    },
    arrow/.style={->, thick},
    node distance=1.65cm  % Decreased from 1.8cm to 1.65cm
]
\node[block, label=above:\textcolor{teal}{\textbf{Input}}] (input) {$\mathbf{x}$};
\node[block, right=1.65cm of input] (W1) {$\mathbf{W}_1$};
\node[dashedblock, above=1.6cm of W1, label=above:\textcolor{orange!70!black}{\textbf{LoRA Contributor}}] (LoRA) {LoRA for Layer 1};
\node[block, right=1.65cm of W1] (act) {ReLU};
\node[block, right=1.65cm of act] (W2) {$\mathbf{W}_2$};
\node[block, label=above:\textcolor{teal}{\textbf{Output}}, right=1.65cm of W2] (output) {$\mathbf{y}$};
\draw[->, teal!60!black] (input) -- node[above, font=\scriptsize]{local multiply} (W1);
\draw[->, dashed, orange!70!black] (W1.north) to[bend left=15] node[left, font=\scriptsize]{partial acts} (LoRA.south);
\draw[->, dashed, orange!70!black] (LoRA.south) to[bend left=15] node[right, font=\scriptsize]{updated $\mathbf{z}'$} (W1.north);
\draw[->, teal!60!black] (W1) -- node[above, font=\scriptsize]{$\mathbf{z}'$} (act);
\draw[->, teal!60!black] (act) -- node[above, font=\scriptsize]{hidden} (W2);
\draw[->, teal!60!black] (W2) -- (output);
\node[draw, rounded corners, thick, color=blue!40!black,
      fit=(input)(W1)(act)(W2)(output),
      label={[font=\small, color=blue!40!black, xshift=2cm]above:\textbf{Base Model User}},
      inner sep=0.6cm
] (baseModelFit) {};
\end{tikzpicture}
\caption{zklora inference flow on a 2-layer neural network where the first layer has private LoRA weights added.}
\label{fig:toy-mlp-lora}
\end{figure}

>>>>

We benchmarked \texttt{ZKLoRA} across several LLMs and smaller models with different numbers of LoRA modules. 
Our central question is how \emph{total} verification, proof generation, and settings times grow with \textbf{the number of LoRA modules} (\#LoRA), while also considering each LoRA's average parameter size. 
Figure~\ref{fig:verify_vs_numlora} and Table~\ref{tab:settings_proof} detail this trade-off.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.9\linewidth]{figs/fig_verify_vs_numlora.pdf}
    \caption{Total verification time (seconds) vs. number of LoRA modules, with dot size reflecting average LoRA size.}
    \label{fig:verify_vs_numlora}
\end{figure}

\begin{table}[h]
\centering
\begin{tabular}{lcccc}
\hline
Base Model & \# of LoRAs & Avg LoRA Params & Avg Settings & Avg Proof \\
\hline
distilgpt2 & 24 & 24576 & 38.0 & 31.6 \\
openai-community/gpt2 & 48 & 49152 & 43.6 & 34.9 \\
meta-llama/Llama-3.2-1B & 32 & 26624 & 37.2 & 31.0 \\
meta-llama/Llama-3.3-70B-Instruct & 80 & 147456 & 54.9 & 46.9 \\
meta-llama/Llama-3.1-8B-Instruct & 32 & 163840 & 57.4 & 47.7 \\
mistralai/Mixtral-8x7B-Instruct-v0.1 & 32 & 327680 & 86.1 & 73.7 \\
\hline
\end{tabular}
\caption{Model Performance Metrics}\label{tab:settings_proof}
\end{table}

@@

Table~\ref{tab:settings_proof} similarly show that \emph{proof generation} and \emph{settings time} scale with the number of LoRA modules, but also depend heavily on each module’s average size (indicated by larger or smaller dot size). These two steps (proof generation on the LoRA contributor’s side and cryptographic circuit setup) can become more expensive, yet remain feasible in decentralized settings or paid contract relationships. The \emph{Base Model User}, meanwhile, benefits from the relatively short verification overhead.

@@

\iffalse
\begin{figure}[ht]
\centering
\begin{tikzpicture}[
    font=\small,
    >=latex,
    thick,
    block/.style={
      draw,
      rounded corners,
      align=center,
      minimum width=1.65cm,  % Decreased from 1.8cm to 1.65cm
      minimum height=0.9cm,
      fill=gray!10  % Light gray background for regular blocks
    },
    dashedblock/.style={
      draw,
      dashed,
      rounded corners,
      align=center,
      font=\footnotesize,
      fill=orange!10,  % Light orange background for LoRA block
      minimum width=1.65cm,  % Decreased to match other blocks
      minimum height=0.9cm
    },
    arrow/.style={->, thick},
    node distance=1.65cm  % Decreased from 1.8cm to 1.65cm
]
\node[block, label=above:\textcolor{teal}{\textbf{Input}}] (input) {$\mathbf{x}$};
\node[block, right=1.65cm of input] (W1) {$\mathbf{W}_1$};
\node[dashedblock, above=1.6cm of W1, label=above:\textcolor{orange!70!black}{\textbf{LoRA Contributor}}] (LoRA) {LoRA for Layer 1};
\node[block, right=1.65cm of W1] (act) {ReLU};
\node[block, right=1.65cm of act] (W2) {$\mathbf{W}_2$};
\node[block, label=above:\textcolor{teal}{\textbf{Output}}, right=1.65cm of W2] (output) {$\mathbf{y}$};
\draw[->, teal!60!black] (input) -- node[above, font=\scriptsize]{local multiply} (W1);
\draw[->, dashed, orange!70!black] (W1.north) to[bend left=15] node[left, font=\scriptsize]{partial acts} (LoRA.south);
\draw[->, dashed, orange!70!black] (LoRA.south) to[bend left=15] node[right, font=\scriptsize]{updated $\mathbf{z}'$} (W1.north);
\draw[->, teal!60!black] (W1) -- node[above, font=\scriptsize]{$\mathbf{z}'$} (act);
\draw[->, teal!60!black] (act) -- node[above, font=\scriptsize]{hidden} (W2);
\draw[->, teal!60!black] (W2) -- (output);
\node[draw, rounded corners, thick, color=blue!40!black,
      fit=(input)(W1)(act)(W2)(output),
      label={[font=\small, color=blue!40!black, xshift=2cm]above:\textbf{Base Model User}},
      inner sep=0.6cm
] (baseModelFit) {};
\end{tikzpicture}
\caption{zklora inference flow on a 2-layer neural network where the first layer has private LoRA weights added.}
\label{fig:toy-mlp-lora}
\end{figure}
\fi

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
    font=\small,
    >=latex,
    thick,
    block/.style={
      draw,
      rounded corners,
      align=center,
      minimum width=2cm,
      minimum height=1.2cm,
      fill=blue!5
    },
    dashedblock/.style={
      draw,
      dashed,
      rounded corners,
      align=center,
      fill=orange!10,
      minimum width=2cm,
      minimum height=1.2cm
    },
    sumblock/.style={
      draw,
      circle,
      fill=gray!20,
      inner sep=1pt,
      minimum size=0.5cm
    },
    arrow/.style={->, thick},
    node distance=2.5cm
]

% 1) Input x
\node[block, label=above:\textcolor{teal}{\textbf{Input}}] (xnode) {\(\mathbf{x}\)};

% 2) Local submodule (removed label)
\node[block, right=2.5cm of xnode] (localSub) {\textbf{Wx}};

% 3) Summation or aggregator (removed z label)
\node[sumblock, right=2.5cm of localSub] (sumNode) {\(+\)};

% 4) Final output
\node[block, right=2.5cm of sumNode, 
      label=above:\textcolor{teal}{\textbf{Final Output}}] (finalOut) {%
\(\textbf{h}\)};

% 5) The dashed remote LoRA block
\node[dashedblock, above=1.8cm of localSub, 
      label=above:\textcolor{orange!70!black}{\textbf{Remote LoRA}}] 
(lora) {\(\Delta = \mathbf{BAx}\)};

% ARROWS
% x -> localSub
\draw[->, teal!70!black] (xnode) -- node[above, font=\scriptsize]{\textbf{local forward}} (localSub);

% localSub -> aggregator
\draw[->, teal!70!black] (localSub) -- 
   node[above, font=\scriptsize]{\(\textbf{base out}\)} (sumNode);

% aggregator -> final
\draw[->, teal!70!black] (sumNode) -- (finalOut);

% x -> remote LoRA (increased font size)
\draw[->, dashed, orange!70!black] (xnode.north) to[bend left=18]
   node[left, font=\small]{send \(\mathbf{x}\)} (lora.west);

% remote LoRA -> aggregator (increased font size)
\draw[->, dashed, orange!70!black] (lora.east) to[bend right=15]
   node[right, font=\small]{\(\mathbf{\Delta}\)} (sumNode.north);

% Base Model box with label shifted left
\node[
  draw,
  rounded corners,
  thick,
  color=blue!40!black,
  fit=(xnode)(localSub)(sumNode)(finalOut),
  label={[font=\small, color=blue!40!black, xshift=-1.5cm]above:\textbf{Base Model User}}
] (baseBox) {};

\end{tikzpicture}
\caption{Flow in a Multi-Party Inference scenario between local base model and remote LoRA weights. 
The base model performs a local forward pass, computing \(\mathbf{base\_out} = \mathbf{Wx}\). In parallel, the input \(\mathbf{x}\) is sent to the remote LoRA module, which returns \(\mathbf{\Delta} = \mathbf{BAx}\). The final output is \(\mathbf{base\_out} + \mathbf{\Delta}\).}
\label{fig:remote-lora-flow}
\end{figure}
